Data_Analysis_MS_ddm_exp1

### 2019-01-03 Results/preproc/MS_ddm_preproc.rmd
Check the results, and start using tidyverse for preprocessing

This script is aimed at making the analysis of the moral self experiment (DDM) reproducible. We analyzezd the data using the traditional ways (d prime/Accuracy and mean of RT)

```{r Initializing, include=FALSE}
source('Initial.r')
curDir  <- getwd()
#curDir  <- dirname(rstudioapi::getSourceEditorContext()$path)   # directory for preprocessing
rootDir <- gsub('.{7}$', '', curDir)
#rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'Traditional_Analysis',sep = '')
ddmDir <- paste(rootDir,'HDDM',sep = '')
exgDir <- paste(rootDir,'exGaussian',sep = '')
```

@@

# make the variables in a specified order
df.C$Morality <- factor(df.C$Morality, levels=c("Good","Bad"))    
df.C$Identity <- factor(df.C$Identity, levels=c("Self","Other"))

@@

# exclude trials, criterio 2: practicing trials in matching task (first 48 trials)
subNo <- unique(df.M$Subject)
for (subj in subNo) {
        if (exists('df.M.fm')){
                df.tmp <- df.M[df.M$Subject == subj,]
                df.tmp <- df.tmp[49:nrow(df.tmp),]
                df.M.fm <- rbind(df.M.fm,df.tmp)
        } else {
                df.M.fm <- df.M[df.M$Subject == subj,]
                df.M.fm <- df.M.fm[49:nrow(df.M.fm),]
        }
}   # df.M.fm should be 14880 rows
df.M <- df.M.fm  # all the experimental trials
rm(subNo,df.M.fm,df.tmp)

# exlude subject criterion 2: less than 50% overall accuracy
# change no response to error
df.M.tmp <- df.M
df.M.tmp$ACC[df.M.tmp$ACC == -1] <- 0
df.C.tmp <- df.C
df.C.tmp$ACC[df.C.tmp$ACC == -1] <- 0 # this was not done in previous analysis, resulting the wronly excluded participants in registration

@@

excldSub2.M <- df.M.acc.g$Subject[df.M.acc.g$ACC < 0.5]      # < 50% accuracy in matching task
excldSub2.C <- df.C.acc.g$Subject[df.C.acc.g$ACC < 0.5]      # < 50% accuracy in categorization task
df.M.valid <- df.M[!(df.M$Subject %in% excldSub2.M),]        # exclude the invalid subjects
df.C.valid <- df.C[!(df.C$Subject %in% excldSub2.M),]
df.C.valid <- df.C.valid[!(df.C.valid$Subject %in% excldSub2.C),]
rm(df.M,df.C,df.M.tmp,df.C.tmp)
```

@@

# excld.trials, criterion 2: correct response within 200 ms
excld.trials.M <- df.M1[df.M1$RT <= 200,]
df.M1.V        <- df.M1[!(df.M1$RT <= 200),]  # valid trial data for match task
excld.trials.C <- df.C1[df.C1$RT <= 200,]
df.C1.V        <- df.C1[!(df.C1$RT <= 200),]  # valid trial data for categorization task
nrow(excld.trials.C) + nrow(df.C1.V) == nrow(df.C1) # if true, everything is ok

## Basic information of the data ####
df.M1.T.basic <- df.M1[!duplicated(df.M1$Subject), 1:4]
numT.subj <- nrow(df.M1.T.basic)
numT.female <- sum(df.M1.T.basic$Gender == 'female');
numT.male <- sum(df.M1.T.basic$Gender == 'male');
ageT.mean <- round(mean(df.M1.T.basic$Age),2);
ageT.std <- round(sd(df.M1.T.basic$Age),2);
num.excldSub2.M <- length(unique(excldSub2.M))
num.excldSub2.C <- length(unique(excldSub2.C))

# valide data for matching task
df.M1.V.basic <- df.M1.V[!duplicated(df.M1.V$Subject), 1:4]
numV.female <- sum(df.M1.V.basic$Gender == 'female');
numV.male <- sum(df.M1.V.basic$Gender == 'male');
ageV.mean <- round(mean(df.M1.V.basic$Age),2);
ageV.std <- round(sd(df.M1.V.basic$Age),2);
ratio.excld.trials.M <- nrow(excld.trials.M)/nrow(df.M1)

# valid data for categorization task
df.C1.V.basic <- df.C1.V[!duplicated(df.C1.V$Subject), 1:4]
numV.female.C <- sum(df.C1.V.basic$Gender == 'female');
numV.male.C <- sum(df.C1.V.basic$Gender == 'male');
ageV.mean.C <- round(mean(df.C1.V.basic$Age),2);
ageV.std.C <- round(sd(df.C1.V.basic$Age),2);
ratio.excld.trials.C <- nrow(excld.trials.C)/nrow(df.C1)

rm(df.M1,df.C1) # remove the no longer used varibles.

# get the data file for hddm analysis ####
df.M.ddm.v <- df.M.valid   # exclude the invalid subjects
df.C.ddm.v <- df.C.valid

# exclusion trials, criterion: trials without response or wrong key
# Note: we didn't remove the correct response less than 200 ms
df.M.hddm <- subset(df.M.ddm.v, ACC ==1 | ACC == 0)  
df.C.hddm <- subset(df.C.ddm.v, ACC ==1 | ACC == 0)

df.M.hddm_m <- subset(df.M.hddm, Match == 'match') # select data
df.M.hddm_m <- df.M.hddm_m[,c('Subject','Morality','Identity','RT','ACC')] # select column
colnames(df.M.hddm_m) <- c("subj_idx","val",'id','rt','response')     # change column name

df.M.hddm_nm <- subset(df.M.hddm, Match == 'nonmatch')
df.M.hddm_nm <- df.M.hddm_nm[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.M.hddm_nm) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_val <- subset(df.C.hddm, Task == 'Val')
df.C.hddm_val <- df.C.hddm_val[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_val) <- c("subj_idx","val",'id','rt','response')

df.C.hddm_id <- subset(df.C.hddm, Task == 'Id')
df.C.hddm_id <- df.C.hddm_id[,c('Subject','Morality','Identity','RT','ACC')]
colnames(df.C.hddm_id) <- c("subj_idx","val",'id','rt','response')

setwd(ddmDir)
write.csv(df.M.hddm_m,'MS_match_hddm.csv',row.names = F)
write.csv(df.M.hddm_nm,'MS_mismatch_hddm.csv',row.names = F)
write.csv(df.C.hddm_val,'MS_categ_val_hddm.csv',row.names = F)
write.csv(df.C.hddm_id,'MS_categ_id_hddm.csv',row.names = F)
setwd(curDir)
rm('df.M.hddm','df.C.hddm','df.M.hddm_m','df.M.hddm_nm','df.C.hddm_val','df.C.hddm_id')

# get the data for exGaussian analysis
df.C.exG <- df.C.valid
df.C.exG <- subset(df.C.exG, ACC == 1 & RT > .2)
setwd(exgDir)
write.csv(df.C.exG,'MS_categ_exG.csv',row.names = F)
rm(df.C.exG)

>>>>

This script is used to preprocessing the data of the moral self experiment (DDM), experiment 1.

Input: 
  MS_matchTask_raw.csv -- raw data of matching task;
  MS_categTask_raw.csv -- raw data of categorization task;

Output:
  df.M.hddm_m.csv  -- clean data of matching trials from matchign task, for HDDM analysis;
  df.M.hddm_nm.csv -- clean data of nonmatching trials from matchign task, for HDDM analysis;
  df.C.hddm_val    -- clean data of valence-based categorization trials, for HDDM analysis;
  df.C.hddm_id     -- clean data of identity-based categorization trials, for HDDM analysis;

  MS_categ_behav_wide.csv -- cleaned summary results (wide-format) of categorization task for statistical analysis in JASP
  MS_categ_behav_noTask_wide.csv -- cleaned summary results (wide-format) of categorization task (collapsed data from different task) for statistical analysis in JASP
  MS_categ__rt_acc_long.csv -- cleaned summary results of RT and accuracy (long-format) of categorization task, for JASP analysis
  MS_categ__rt_acc_noTask_long.csv -- cleaned summary reuslt of RT and accuracy (long-format) of categorization task (collapsed data from different task) for JASP analysis
  MS_cross_taskeffect_wide.csv-- cleaned summary results for cross task analysis

```{r Initializing, include=FALSE}
source('Initial.r')
curDir  <- getwd() # get the current working directory
#curDir  <- dirname(rstudioapi::getSourceEditorContext()$path)   # directory for preprocessing

rootDir <- gsub('.{7}$', '', curDir)                      # get the parental folder
#rawDir <- paste(curDir,'/data/',sep = '')
traDir <- paste(rootDir,'Traditional_Analysis',sep = '')  # folder for traditional analsysi
ddmDir <- paste(rootDir,'HDDM',sep = '')                  # folder for DDM analysis
exgDir <- paste(rootDir,'exGaussian',sep = '')            # folder for ExGaussian analysis
```

@@

df.C$Morality <- factor(df.C$Morality, levels=c("Good","Bad"))    
df.C$Identity <- factor(df.C$Identity, levels=c("Self","Other"))

@@

# exclude trials, criterio 2: practicing trials in matching task (first 48 trials)
subNo <- unique(df.M$Subject)
for (subj in subNo) {
        if (exists('df.M.fm')){
                df.tmp <- df.M[df.M$Subject == subj,]
                df.tmp <- df.tmp[49:nrow(df.tmp),]
                df.M.fm <- rbind(df.M.fm,df.tmp)
        } else {
                df.M.fm <- df.M[df.M$Subject == subj,]
                df.M.fm <- df.M.fm[49:nrow(df.M.fm),]
        }
}   # df.M.fm should be 14880 rows
df.M <- df.M.fm  # all the experimental trials
rm(subNo,df.M.fm,df.tmp) # remove the intermedia variables

# exlude subject criterion 2: less than 50% overall accuracy
# re-code trials without response as incorrect
df.M.tmp <- df.M
df.M.tmp$ACC[df.M.tmp$ACC == -1] <- 0
df.C.tmp <- df.C
df.C.tmp$ACC[df.C.tmp$ACC == -1] <- 0 # this was not done in previous analysis, resulting the wronly excluded participants in registration

@@

excldSub2.M <- df.M.acc.g$Subject[df.M.acc.g$ACC < 0.5]      # < 50% accuracy in matching task
excldSub2.C <- df.C.acc.g$Subject[df.C.acc.g$ACC < 0.5]      # < 50% accuracy in categorization task
df.M.valid <- df.M[!(df.M$Subject %in% excldSub2.M),]        # exclude the invalid subjects
df.C.valid <- df.C[!(df.C$Subject %in% excldSub2.M),]
df.C.valid <- df.C.valid[!(df.C.valid$Subject %in% excldSub2.C),] # one participants were excluded for categorization task
rm(df.M,df.C,df.M.tmp,df.C.tmp)
```

@@

# excld.trials, criterion 3: correct response within 200 ms
excld.trials.M <- df.M1[df.M1$RT <= 200,]
df.M1.V        <- df.M1[!(df.M1$RT <= 200),]        # valid trial data for match task
excld.trials.C <- df.C1[df.C1$RT <= 200,]
df.C1.V        <- df.C1[!(df.C1$RT <= 200),]        # valid trial data for categorization task
nrow(excld.trials.C) + nrow(df.C1.V) == nrow(df.C1) # if true, everything is ok

## Basic information of the data ####
df.M1.T.basic.info <- df.M1 %>%
  dplyr::select(Subject, Age, Gender) %>%
  dplyr::distinct(Subject, Age, Gender) %>%
  dplyr::summarise(N = length(Subject),
                   N.f = sum(Gender == 'female'),
                   N.m = sum(Gender == 'male'),
                   MeanAge = round(mean(Age),2),
                   SDAge   = round(sd(Age),2))
#numT.subj <- nrow(df.M1.T.basic)
#numT.female <- sum(df.M1.T.basic$Gender == 'female');
#numT.male <- sum(df.M1.T.basic$Gender == 'male');
#ageT.mean <- round(mean(df.M1.T.basic$Age),2);
#ageT.std <- round(sd(df.M1.T.basic$Age),2);
num.excldSub2.M <- length(unique(excldSub2.M))
num.excldSub2.C <- length(unique(excldSub2.C))

# valid data for matching task
df.M1.V.basic.info <- df.M1.V %>%
  dplyr::select(Subject, Age, Gender) %>%
  dplyr::distinct(Subject, Age, Gender) %>%
  dplyr::summarise(N = length(Subject),
                   N.f = sum(Gender == 'female'),
                   N.m = sum(Gender == 'male'),
                   MeanAge = round(mean(Age),2),
                   SDAge   = round(sd(Age),2))

ratio.excld.trials.M <- nrow(excld.trials.M)/nrow(df.M1)

# valid data for categorization task
df.C1.V.basic.info <- df.C1.V %>%
  dplyr::select(Subject, Age, Gender) %>%
  dplyr::distinct(Subject, Age, Gender) %>%
  dplyr::summarise(N = length(Subject),
                   N.f = sum(Gender == 'female'),
                   N.m = sum(Gender == 'male'),
                   MeanAge = round(mean(Age),2),
                   SDAge   = round(sd(Age),2))

ratio.excld.trials.C <- nrow(excld.trials.C)/nrow(df.C1)

###########################################################
########   get the data file for hddm analysis      #######
###########################################################
# exclusion trials, criterion: trials without response or wrong key
# Note: we didn't remove the correct response less than 200 ms

df.M.hddm_m <- df.M.valid %>%                          # start from the valid data, RT in seconds.
  dplyr::filter(ACC == 1 | ACC == 0) %>%               # exclude trials without response or with wrong keys
  dplyr::filter(Match == 'match') %>%
  dplyr::select(Subject,Morality,Identity,RT,ACC) %>%  # select column
  dplyr::rename(subj_idx = Subject, val = Morality, id = Identity, rt = RT, response = ACC) # rename column

df.M.hddm_nm <- df.M.valid %>%
  dplyr::filter(ACC == 1 | ACC == 0) %>%               # exclude trials without response or with wrong keys
  dplyr::filter(Match == 'nonmatch') %>%
  dplyr::select(Subject,Morality,Identity,RT,ACC) %>%  # select column
  dplyr::rename(subj_idx = Subject, val = Morality, id = Identity, rt = RT, response = ACC) # rename column

df.C.hddm_val <- df.C.valid %>%
  dplyr::filter(ACC == 1 | ACC == 0) %>%               # exclude trials without response or with wrong keys
  dplyr::filter(Task == 'Val') %>%                     # select the valence-based task
  dplyr::select(Subject,Morality,Identity,RT,ACC) %>%  # select column
  dplyr::rename(subj_idx = Subject, val = Morality, id = Identity, rt = RT, response = ACC) # rename column

df.C.hddm_id <- df.C.valid %>%
  dplyr::filter(ACC == 1 | ACC == 0) %>%               # exclude trials without response or with wrong keys
  dplyr::filter(Task == 'Id') %>%                      # select the valence-based task
  dplyr::select(Subject,Morality,Identity,RT,ACC) %>%  # select column
  dplyr::rename(subj_idx = Subject, val = Morality, id = Identity, rt = RT, response = ACC) # rename column

setwd(ddmDir)
write.csv(df.M.hddm_m,'MS_match_hddm.csv',row.names = F)
write.csv(df.M.hddm_nm,'MS_mismatch_hddm.csv',row.names = F)
write.csv(df.C.hddm_val,'MS_categ_val_hddm.csv',row.names = F)
write.csv(df.C.hddm_id,'MS_categ_id_hddm.csv',row.names = F)
setwd(curDir)
rm('df.M.hddm_m','df.M.hddm_nm','df.C.hddm_val','df.C.hddm_id')

###########################################################
######## get the data file for exGaussian analysis  #######
###########################################################
df.C.exG <- df.C.valid %>%
  dplyr::filter(ACC == 1 & RT > .2)    # only for correct and valid trials

setwd(exgDir)
write.csv(df.C.exG,'MS_categ_exG.csv',row.names = F)
rm(df.C.exG)
